---
title: "Anchor Representation: 2024 vs. 2026"
format: pdf
editor: visual
---
```{r, echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)

suppressPackageStartupMessages({
  library(tidyverse)
})
```

# Overview

This report compares anchor representation between the 2024 and 2026 calibration datasets.

The purpose is to determine which anchors were sufficiently represented in each calibration period and which anchors were excluded because they did not meet the representation criterion.

For **both 2024 and 2026**, an anchor is considered sufficiently represented if it appears in at least **80% of the available calibration files**.

This analysis distinguishes between:

* anchors retained in both analyses;
* anchors retained in 2024 but not 2026;
* anchors retained in 2026 but not 2024;
* anchors that appeared in the raw data but were excluded from both analyses.

The analysis uses anchor IDs only. Calibration values are not required for this comparison.

---

# 1. Load 2024 Calibration Files

The 2024 files are identified by their filename pattern.

```{r}
files_2024 <- list.files(
  "calibration_outputs",
  pattern = "^calibrations_probeData_2024-\\d{2}-\\d{2}\\.csv$",
  full.names = TRUE
) %>%
  sort()

length(files_2024)
```

The number above is the total number of available 2024 calibration files.

---

# 2. Read 2024 Anchor IDs

Each 2024 file is read only to obtain the anchor IDs.

The IDs are immediately converted to character so that they have the same type used throughout the 2026 analysis.

```{r}
raw_ids_2024 <- map_dfr(
  seq_along(files_2024),
  function(i) {

    dat <- read_csv(
      files_2024[i],
      col_names = FALSE,
      col_types = cols(.default = col_character()),
      na = c("", "NA")
    )

    tibble(
      id = as.character(dat[[1]]),
      file = basename(files_2024[i])
    )
  }
)
```

Remove missing IDs:

```{r}
raw_ids_2024 <- raw_ids_2024 %>%
  filter(
    !is.na(id),
    id != ""
  ) %>%
  mutate(
    id = as.character(id)
  )
```

---

# 3. Apply the 80% Representation Criterion to 2024

An anchor must appear in at least 80% of the available 2024 calibration files.

```{r}
n_files_2024 <- length(files_2024)

threshold_2024 <- 0.80 * n_files_2024

threshold_2024
```

Count the number of files containing each anchor:

```{r}
anchor_counts_2024 <- raw_ids_2024 %>%
  distinct(
    id,
    file
  ) %>%
  count(
    id,
    name = "files_present"
  ) %>%
  mutate(
    total_files = n_files_2024,
    proportion_present = files_present / total_files,
    retained_80pct = files_present >= threshold_2024
  ) %>%
  arrange(
    desc(files_present),
    id
  )
```

The anchors retained for the 2024 analysis are:

```{r}
retained_2024 <- anchor_counts_2024 %>%
  filter(retained_80pct) %>%
  select(id) %>%
  distinct() %>%
  mutate(
    id = as.character(id)
  )
```

Number of retained 2024 anchors:

```{r}
nrow(retained_2024)
```

---

# 4. Load 2026 Calibration Files

```{r}
files_2026 <- list.files(
  "26_Calibration",
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  sort()

length(files_2026)
```

---

# 5. Remove 2026 Files with Fewer Than 500 Observations

This reproduces the filtering used in the primary 2026 calibration analysis.

```{r}
df_list_2026 <- lapply(
  files_2026,
  read.csv
)

keep_2026 <- sapply(
  df_list_2026,
  function(x) {
    !is.null(x) &&
      nrow(x) >= 500
  }
)

df_list_2026 <- df_list_2026[keep_2026]

files_2026_retained <- files_2026[keep_2026]

n_files_2026 <- length(files_2026_retained)

n_files_2026
```

The 2026 representation analysis therefore uses only files containing at least 500 observations, consistent with the primary 2026 analysis.

---

# 6. Read 2026 Anchor IDs

```{r}
raw_ids_2026 <- map_dfr(
  seq_along(df_list_2026),
  function(i) {

    dat <- df_list_2026[[i]]

    tibble(
      id = as.character(dat[[1]]),
      file = basename(files_2026_retained[i])
    )
  }
)
```

Remove missing IDs:

```{r}
raw_ids_2026 <- raw_ids_2026 %>%
  filter(
    !is.na(id),
    id != ""
  ) %>%
  mutate(
    id = as.character(id)
  )
```

---

# 7. Apply the 80% Representation Criterion to 2026

An anchor must appear in at least 80% of the usable 2026 calibration files.

```{r}
threshold_2026 <- 0.80 * n_files_2026

threshold_2026
```

Count files containing each anchor:

```{r}
anchor_counts_2026 <- raw_ids_2026 %>%
  distinct(
    id,
    file
  ) %>%
  count(
    id,
    name = "files_present"
  ) %>%
  mutate(
    total_files = n_files_2026,
    proportion_present = files_present / total_files,
    retained_80pct = files_present >= threshold_2026
  ) %>%
  arrange(
    desc(files_present),
    id
  )
```

The anchors retained for the 2026 analysis are:

```{r}
retained_2026 <- anchor_counts_2026 %>%
  filter(retained_80pct) %>%
  select(id) %>%
  distinct() %>%
  mutate(
    id = as.character(id)
  )
```

Number of retained 2026 anchors:

```{r}
nrow(retained_2026)
```

---

# 8. Verify ID Types

Before comparing the datasets, explicitly force both ID columns to character.

```{r}
retained_2024 <- retained_2024 %>%
  mutate(id = as.character(id))

retained_2026 <- retained_2026 %>%
  mutate(id = as.character(id))

raw_ids_2024 <- raw_ids_2024 %>%
  mutate(id = as.character(id))

raw_ids_2026 <- raw_ids_2026 %>%
  mutate(id = as.character(id))
```

Verify:

```{r}
data.frame(
  dataset = c(
    "retained_2024",
    "retained_2026",
    "raw_ids_2024",
    "raw_ids_2026"
  ),
  id_type = c(
    class(retained_2024$id)[1],
    class(retained_2026$id)[1],
    class(raw_ids_2024$id)[1],
    class(raw_ids_2026$id)[1]
  )
)
```

All four should report `character`.

---

# 9. Anchors Retained in Both Years

These are anchors that meet the 80% criterion in both datasets.

```{r}
retained_both <- inner_join(
  retained_2024,
  retained_2026,
  by = "id"
) %>%
  arrange(id)
```

Number retained in both:

```{r}
nrow(retained_both)
```

View the IDs:

```{r}
retained_both
```

---

# 10. Anchors Retained in 2024 but Not 2026

These anchors met the 80% criterion in 2024 but did not meet it in 2026.

```{r}
retained_2024_only <- anti_join(
  retained_2024,
  retained_2026,
  by = "id"
) %>%
  arrange(id)
```

Number:

```{r}
nrow(retained_2024_only)
```

IDs:

```{r}
retained_2024_only
```

---

# 11. Anchors Retained in 2026 but Not 2024

These anchors met the 80% criterion in 2026 but did not meet it in 2024.

```{r}
retained_2026_only <- anti_join(
  retained_2026,
  retained_2024,
  by = "id"
) %>%
  arrange(id)
```

Number:

```{r}
nrow(retained_2026_only)
```

IDs:

```{r}
retained_2026_only
```

---

# 12. All Raw Anchors Observed in 2024

This represents every unique anchor that appeared at least once in the original 2024 calibration files.

```{r}
all_raw_2024 <- raw_ids_2024 %>%
  distinct(id) %>%
  mutate(
    id = as.character(id)
  ) %>%
  arrange(id)
```

Number:

```{r}
nrow(all_raw_2024)
```

---

# 13. All Raw Anchors Observed in 2026

This represents every unique anchor that appeared at least once in the usable 2026 calibration files.

```{r}
all_raw_2026 <- raw_ids_2026 %>%
  distinct(id) %>%
  mutate(
    id = as.character(id)
  ) %>%
  arrange(id)
```

Number:

```{r}
nrow(all_raw_2026)
```

---

# 14. Anchors Excluded from Both Analyses

An anchor is considered excluded from both analyses if:

* it appeared in the raw 2024 and/or 2026 data;
* it did not meet the 80% criterion in 2024;
* it did not meet the 80% criterion in 2026.

First create the complete set of anchors observed in either year.

```{r}
all_raw_anchors <- union(
  all_raw_2024$id,
  all_raw_2026$id
) %>%
  as.character()
```

Then remove every anchor retained in either analysis:

```{r}
retained_either_year <- union(
  retained_2024$id,
  retained_2026$id
) %>%
  as.character()
```

The resulting set consists of anchors excluded from both analyses:

```{r}
excluded_both <- tibble(
  id = setdiff(
    all_raw_anchors,
    retained_either_year
  )
) %>%
  arrange(id)
```

Number excluded from both:

```{r}
nrow(excluded_both)
```

IDs:

```{r}
excluded_both
```

---

# 15. Determine Where the Excluded Anchors Appeared

It is useful to distinguish anchors that were excluded from both but appeared in:

* 2024 only;
* 2026 only;
* both years.

## 15.1 Excluded from Both, Present in Both Raw Datasets

```{r}
excluded_both_raw <- excluded_both %>%
  filter(
    id %in% all_raw_2024$id,
    id %in% all_raw_2026$id
  ) %>%
  arrange(id)
```

Number:

```{r}
nrow(excluded_both_raw)
```

IDs:

```{r}
excluded_both_raw
```

---

## 15.2 Excluded from Both, Present in 2024 Only

```{r}
excluded_2024_only_raw <- excluded_both %>%
  filter(
    id %in% all_raw_2024$id,
    !id %in% all_raw_2026$id
  ) %>%
  arrange(id)
```

Number:

```{r}
nrow(excluded_2024_only_raw)
```

IDs:

```{r}
excluded_2024_only_raw
```

---

## 15.3 Excluded from Both, Present in 2026 Only

```{r}
excluded_2026_only_raw <- excluded_both %>%
  filter(
    id %in% all_raw_2026$id,
    !id %in% all_raw_2024$id
  ) %>%
  arrange(id)
```

Number:

```{r}
nrow(excluded_2026_only_raw)
```

IDs:

```{r}
excluded_2026_only_raw
```

---

# 16. Detailed Representation Table

The following table combines the representation information for every anchor observed in either year.

```{r}
representation_comparison <- full_join(
  anchor_counts_2024 %>%
    select(
      id,
      files_2024 = files_present,
      total_files_2024 = total_files,
      proportion_2024 = proportion_present,
      retained_2024 = retained_80pct
    ),

  anchor_counts_2026 %>%
    select(
      id,
      files_2026 = files_present,
      total_files_2026 = total_files,
      proportion_2026 = proportion_present,
      retained_2026 = retained_80pct
    ),

  by = "id"
) %>%
  mutate(
    id = as.character(id),

    present_2024 = !is.na(files_2024),
    present_2026 = !is.na(files_2026),

    retained_2024 = replace_na(
      retained_2024,
      FALSE
    ),

    retained_2026 = replace_na(
      retained_2026,
      FALSE
    ),

    classification = case_when(

      retained_2024 & retained_2026 ~
        "Retained in both",

      retained_2024 & !retained_2026 ~
        "Retained in 2024 only",

      !retained_2024 & retained_2026 ~
        "Retained in 2026 only",

      !retained_2024 & !retained_2026 &
        present_2024 & present_2026 ~
        "Excluded from both; present in both raw datasets",

      !retained_2024 & !retained_2026 &
        present_2024 & !present_2026 ~
        "Excluded from both; 2024 only",

      !retained_2024 & !retained_2026 &
        !present_2024 & present_2026 ~
        "Excluded from both; 2026 only",

      TRUE ~
        "Other"
    )
  ) %>%
  arrange(
    classification,
    id
  )
```

View the complete table:

```{r}
representation_comparison
```

---

# 17. Summary of Anchor Classification

```{r}
classification_summary <- representation_comparison %>%
  count(
    classification,
    name = "number_of_anchors"
  ) %>%
  arrange(
    desc(number_of_anchors)
  )

classification_summary
```

---

# 18. Summary Table for the Report

A more compact summary is useful for reporting the results.

```{r}
final_summary <- tibble(
  category = c(
    "Retained in both 2024 and 2026",
    "Retained in 2024 only",
    "Retained in 2026 only",
    "Excluded from both; present in both raw datasets",
    "Excluded from both; present in 2024 only",
    "Excluded from both; present in 2026 only"
  ),

  number_of_anchors = c(
    nrow(retained_both),
    nrow(retained_2024_only),
    nrow(retained_2026_only),
    nrow(excluded_both_raw),
    nrow(excluded_2024_only_raw),
    nrow(excluded_2026_only_raw)
  )
)

final_summary
```

---

# 19. Representation Percentages

The following calculates the percentage of all anchors observed in either raw dataset that fall into each category.

```{r}
final_summary_percentage <- final_summary %>%
  mutate(
    percentage_of_all_observed =
      number_of_anchors /
      length(all_raw_anchors)
  )

final_summary_percentage
```

---

# 20. Compare 80% Representation Rates

This table summarizes the number of anchors observed in each year and the number that passed the 80% criterion.

```{r}
representation_rates <- tibble(

  year = c(
    "2024",
    "2026"
  ),

  total_files = c(
    n_files_2024,
    n_files_2026
  ),

  eighty_percent_threshold =
    c(
      ceiling(threshold_2024),
      ceiling(threshold_2026)
    ),

  raw_anchors = c(
    nrow(all_raw_2024),
    nrow(all_raw_2026)
  ),

  retained_80pct = c(
    nrow(retained_2024),
    nrow(retained_2026)
  )
) %>%
  mutate(
    proportion_retained =
      retained_80pct / raw_anchors
  )

representation_rates
```

---

# 21. Anchors Retained in One Year but Not the Other

This table is particularly useful for identifying changes in anchor availability between calibration campaigns.

```{r}
year_specific_anchors <- bind_rows(

  retained_2024_only %>%
    mutate(
      classification = "Retained in 2024 only"
    ),

  retained_2026_only %>%
    mutate(
      classification = "Retained in 2026 only"
    )
) %>%
  arrange(
    classification,
    id
  )

year_specific_anchors
```

---

# 22. Anchors Excluded from Both but Observed in Either Dataset

This is the group of anchors that may be particularly useful for investigating why anchors failed the representation criterion.

```{r}
excluded_from_both_detailed <- representation_comparison %>%
  filter(
    classification %in% c(
      "Excluded from both; present in both raw datasets",
      "Excluded from both; 2024 only",
      "Excluded from both; 2026 only"
    )
  ) %>%
  select(
    id,
    files_2024,
    total_files_2024,
    proportion_2024,
    files_2026,
    total_files_2026,
    proportion_2026,
    classification
  ) %>%
  arrange(
    classification,
    id
  )

excluded_from_both_detailed
```

---

# 23. Most Poorly Represented Anchors

The following table identifies anchors with the lowest representation in either year.

This can help identify anchors that were particularly intermittent.

```{r}
poorly_represented <- representation_comparison %>%
  mutate(
    representation_2024 =
      replace_na(
        proportion_2024,
        0
      ),

    representation_2026 =
      replace_na(
        proportion_2026,
        0
      ),

    minimum_representation =
      pmin(
        representation_2024,
        representation_2026
      )
  ) %>%
  select(
    id,
    representation_2024,
    representation_2026,
    minimum_representation,
    classification
  ) %>%
  arrange(
    minimum_representation,
    id
  )

head(
  poorly_represented,
  50
)
```

---

# 24. Export Results

The major results are written to CSV files so they can be examined outside of the PDF report.

```{r}
write_csv(
  representation_comparison,
  "anchor_representation_comparison_all.csv"
)

write_csv(
  retained_both,
  "anchors_retained_both_2024_2026.csv"
)

write_csv(
  retained_2024_only,
  "anchors_retained_2024_only.csv"
)

write_csv(
  retained_2026_only,
  "anchors_retained_2026_only.csv"
)

write_csv(
  excluded_both,
  "anchors_excluded_both.csv"
)

write_csv(
  excluded_both_raw,
  "anchors_excluded_both_present_both_years.csv"
)

write_csv(
  excluded_2024_only_raw,
  "anchors_excluded_both_2024_only.csv"
)

write_csv(
  excluded_2026_only_raw,
  "anchors_excluded_both_2026_only.csv"
)
```

---

# 25. Final Interpretation

The 80% representation criterion is applied independently to the two calibration periods.

An anchor retained in 2024 therefore appeared in at least 80% of the available 2024 calibration files, while an anchor retained in 2026 appeared in at least 80% of the usable 2026 calibration files.

The comparison separates three scientifically distinct groups:

1. **Stable/repeated anchors:** retained in both 2024 and 2026.
2. **Year-specific anchors:** retained in one calibration period but not the other.
3. **Intermittent anchors:** observed in the raw data but excluded because they failed the 80% representation criterion.

The final group can be particularly informative because failure to meet the 80% criterion may itself reflect changes in anchor availability, connectivity, geographic coverage, or other properties of the calibration system.

```{r}
final_summary
```

```

### One important detail

This version applies the criterion symmetrically, but **the denominator is not necessarily the same between years**. For 2024 it is the number of 2024 files found; for 2026 it is the number of 2026 files remaining after the `<500 observations` filter.

So, for example, if there are 100 usable 2024 files, an anchor needs at least 80 appearances. If there are 80 usable 2026 files, it needs at least 64 appearances.

That is the appropriate interpretation of "present in at least 80% of the available files" and keeps the criterion independent of the different amounts of missing data in the two campaigns.

The resulting **`representation_comparison`** table is probably the most valuable object produced by this QMD: it gives you, for every anchor, its 2024 representation rate, 2026 representation rate, whether it passed each criterion, and its final classification. :contentReference[oaicite:1]{index=1}
```

[1]: https://dplyr.tidyverse.org/reference/filter-joins.html?utm_source=chatgpt.com "Filtering joins — filter-joins • dplyr"
