---
title: "Anchor Representation: 2024 vs. 2026"
format: pdf
editor: visual
author: "Tomoko Hida"
---
```{r, echo=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(kableExtra)
})
```

# Overview
This report compares anchor representation between the 2024 and 2026 calibration datasets.

The purpose is to determine which anchors were sufficiently represented in each calibration period and which anchors were excluded because they did not meet the representation criterion.

For both 2024 and 2026, an anchor is considered sufficiently represented if it appears in at least 80% of the available calibration files.

This analysis distinguishes between:

* anchors retained in both analyses
* anchors retained in 2024 but not 2026
* anchors retained in 2026 but not 2024
* anchors that appeared in the raw data but were excluded from both analyses

\newpage

# 1. 2024 Calibration Files

## 1.1 Load 2024 Files

```{r}
files_2024 <- list.files(
  "calibration_outputs",
  pattern = "^calibrations_probeData_2024-\\d{2}-\\d{2}\\.csv$",
  full.names = TRUE
) %>%
  sort()

len24 <- length(files_2024)
```

* Total number of available 2024 calibration files: `r len24`

## 1.2 Read 2024 Anchor IDs
* IDs are immediately converted to char to be consistent with 2026 data

```{r}
raw_ids_2024 <- map_dfr(
  seq_along(files_2024),
  function(i) {

    dat <- read_csv(
      files_2024[i],
      col_names = FALSE,
      col_types = cols(.default = col_character()),
      na = c("", "NA")
    )

    tibble(
      id = as.character(dat[[1]]),
      file = basename(files_2024[i])
    )
  }
)
```

```{r}
# remove missing IDs
raw_ids_2024 <- raw_ids_2024 %>%
  filter(
    !is.na(id),
    id != ""
  ) %>%
  mutate(
    id = as.character(id)
  )
```

## 1.3 Apply the 80% Representation Criterion to 2024 Data

```{r}
n_files_2024 <- length(files_2024)
threshold_2024 <- 0.80 * n_files_2024
```

```{r}
# Count the number of files containing each anchor
anchor_counts_2024 <- raw_ids_2024 %>%
  distinct(
    id,
    file
  ) %>%
  count(
    id,
    name = "files_present"
  ) %>%
  mutate(
    total_files = n_files_2024,
    proportion_present = files_present / total_files,
    retained_80pct = files_present >= threshold_2024
  ) %>%
  arrange(
    desc(files_present),
    id
  )
```

```{r}
# anchors retained for the 2024 analysis

retained_2024 <- anchor_counts_2024 %>%
  filter(retained_80pct) %>%
  select(id) %>%
  distinct() %>%
  mutate(
    id = as.character(id)
  )

numleft24 <- nrow(retained_2024)
```
* ID must appear in at least 80% of the available 2024 calibration files: `r threshold_2024` of `r len24` files

* Number of retained 2024 anchors: `r numleft24`

# 2. 2026 Calibration Files

```{r}
files_2026 <- list.files(
  "26_Calibration",
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  sort()

len26 <- length(files_2026)
```

## 2.1 Load 2026 Files

* Total number of available 2026 calibration files: `r len26`

We impose a filtering technique to ensure robustness for our comparatively wider period of calibration data collection for 2026. A number of datasets of the 2026 calibration data contain significantly few IDs, and would inhibit our analysis by presenting many missing data points. The 2026 representation analysis therefore uses only files containing at least 500 IDs.

```{r}
df_list_2026 <- lapply(
  files_2026,
  read.csv
)

keep_2026 <- sapply(
  df_list_2026,
  function(x) {
    !is.null(x) &&
      nrow(x) >= 500
  }
)

df_list_2026 <- df_list_2026[keep_2026]

files_2026_retained <- files_2026[keep_2026]

n_files_2026 <- length(files_2026_retained)
```
* Number of 2026 files retained: `r n_files_2026`

## 2.2 Read 2026 Anchor IDs

```{r}
raw_ids_2026 <- map_dfr(
  seq_along(df_list_2026),
  function(i) {

    dat <- df_list_2026[[i]]

    tibble(
      id = as.character(dat[[1]]),
      file = basename(files_2026_retained[i])
    )
  }
)
```

```{r}
# Remove missing IDs:
raw_ids_2026 <- raw_ids_2026 %>%
  filter(
    !is.na(id),
    id != ""
  ) %>%
  mutate(
    id = as.character(id)
  )
```

## 2.3 Apply the 80% Representation Criterion to 2026 Data

```{r}
threshold_2026 <- 0.80 * n_files_2026
```

```{r}
# Count files containing each anchor:
anchor_counts_2026 <- raw_ids_2026 %>%
  distinct(
    id,
    file
  ) %>%
  count(
    id,
    name = "files_present"
  ) %>%
  mutate(
    total_files = n_files_2026,
    proportion_present = files_present / total_files,
    retained_80pct = files_present >= threshold_2026
  ) %>%
  arrange(
    desc(files_present),
    id
  )
```

```{r}
# The anchors retained for the 2026 analysis
retained_2026 <- anchor_counts_2026 %>%
  filter(retained_80pct) %>%
  select(id) %>%
  distinct() %>%
  mutate(
    id = as.character(id)
  )

numleft26 <- nrow(retained_2026)
```

* ID must appear in at least 80% of the usable 2026 calibration files: `r threshold_2026` of `r n_files_2026` files
* Number of retained 2026 anchors: `r numleft26`

\newpage

# 3. Anchor Retention

## 3.1 Anchors Retained in Both Years

```{r}
# Before comparing the datasets, ensure both ID columns are type character
retained_2024 <- retained_2024 %>%
  mutate(id = as.character(id))

retained_2026 <- retained_2026 %>%
  mutate(id = as.character(id))

raw_ids_2024 <- raw_ids_2024 %>%
  mutate(id = as.character(id))

raw_ids_2026 <- raw_ids_2026 %>%
  mutate(id = as.character(id))
```

```{r}
## Verify:
# data.frame(
#   dataset = c(
#     "retained_2024",
#     "retained_2026",
#     "raw_ids_2024",
#     "raw_ids_2026"
#   ),
#   id_type = c(
#     class(retained_2024$id)[1],
#     class(retained_2026$id)[1],
#     class(raw_ids_2024$id)[1],
#     class(raw_ids_2026$id)[1]
#   )
# )
## All four should report `character`
```

```{r}
# anchors that meet the 80% criterion in both datasets
retained_both <- inner_join(
  retained_2024,
  retained_2026,
  by = "id"
) %>%
  arrange(id)
# Number retained in both:
 numretain_2426 <-nrow(retained_both)
# View the IDs:
# retained_both
```
Number retained in both: `r numretain_2426`

## 3.2 Anchors Retained in 2024 but Not 2026

```{r}
retained_2024_only <- anti_join(
  retained_2024,
  retained_2026,
  by = "id"
) %>%
  arrange(id)
# number
n24only <- nrow(retained_2024_only)
# IDs:
# retained_2024_only
```
`r n24only` anchors met the 80% criterion in 2024 but did not meet it in 2026.

## 3.3 Anchors Retained in 2026 but Not 2024

```{r}
retained_2026_only <- anti_join(
  retained_2026,
  retained_2024,
  by = "id"
) %>%
  arrange(id)
# num
n26only <- nrow(retained_2026_only)
# IDs:
# retained_2026_only
```

`r n26only` met the 80% criterion in 2026 but did not meet it in 2024.

## 3.4 All Raw Anchors Observed in 2024

```{r}
all_raw_2024 <- raw_ids_2024 %>%
  distinct(id) %>%
  mutate(
    id = as.character(id)
  ) %>%
  arrange(id)
unique24all <- nrow(all_raw_2024)
```

`r unique24all` unique anchors appeared at least once in the original 2024 calibration files.

## 3.5 All Raw Anchors Observed in 2026

```{r}
all_raw_2026 <- raw_ids_2026 %>%
  distinct(id) %>%
  mutate(
    id = as.character(id)
  ) %>%
  arrange(id)
unique26all <- nrow(all_raw_2026)
```

`r unique26all` unique anchors appeared at least once in the usable 2026 calibration files.

## 3.6 Anchors Excluded from Both Analyses

An anchor is considered excluded from both analyses if:

* it appeared in the raw 2024 and/or 2026 data
* it did not meet the 80% criterion in 2024
* it did not meet the 80% criterion in 2026

```{r}
# complete set of anchors observed in either year
all_raw_anchors <- union(
  all_raw_2024$id,
  all_raw_2026$id
) %>%
  as.character()

# remove every anchor retained in either 24 or 26
retained_either_year <- union(
  retained_2024$id,
  retained_2026$id
) %>%
  as.character()

# result set is anchors excluded from both 24 and 26
excluded_both <- tibble(
  id = setdiff(
    all_raw_anchors,
    retained_either_year
  )
) %>%
  arrange(id)

nexclude2426 <- nrow(excluded_both)
```

Unique anchors excluded from both: `r nexclude2426`

\newpage

# 4. Anchor Representation Summaries

## 4.2 Anchor Retention Results

```{r}
# excluded from both, present in both raw datasets
excluded_both_raw <- excluded_both %>%
  filter(
    id %in% all_raw_2024$id,
    id %in% all_raw_2026$id
  ) %>%
  arrange(id)

# nexcluderaw2426 <- nrow(excluded_both_raw)
```

```{r}
# excluded from both, in 2024 only
excluded_2024_only_raw <- excluded_both %>%
  filter(
    id %in% all_raw_2024$id,
    !id %in% all_raw_2026$id
  ) %>%
  arrange(id)
# nexcluderaw24 <- nrow(excluded_2024_only_raw)
```

```{r}
# excluded from both, in 2026 only
excluded_2026_only_raw <- excluded_both %>%
  filter(
    id %in% all_raw_2026$id,
    !id %in% all_raw_2024$id
  ) %>%
  arrange(id)
# nexcluderaw26 <- nrow(excluded_2026_only_raw)
```

```{r}
# this table combines information for every anchor observed in either 24 or 26
representation_comparison <- full_join(
  anchor_counts_2024 %>%
    select(
      id,
      files_2024 = files_present,
      total_files_2024 = total_files,
      proportion_2024 = proportion_present,
      retained_2024 = retained_80pct
    ),

  anchor_counts_2026 %>%
    select(
      id,
      files_2026 = files_present,
      total_files_2026 = total_files,
      proportion_2026 = proportion_present,
      retained_2026 = retained_80pct
    ),

  by = "id"
) %>%
  mutate(
    id = as.character(id),

    present_2024 = !is.na(files_2024),
    present_2026 = !is.na(files_2026),

    retained_2024 = replace_na(
      retained_2024,
      FALSE
    ),

    retained_2026 = replace_na(
      retained_2026,
      FALSE
    ),

    classification = case_when(

      retained_2024 & retained_2026 ~
        "Retained in both",

      retained_2024 & !retained_2026 ~
        "Retained in 2024 only",

      !retained_2024 & retained_2026 ~
        "Retained in 2026 only",

      !retained_2024 & !retained_2026 &
        present_2024 & present_2026 ~
        "Excluded from both; present in both raw datasets",

      !retained_2024 & !retained_2026 &
        present_2024 & !present_2026 ~
        "Excluded from both; 2024 only",

      !retained_2024 & !retained_2026 &
        !present_2024 & present_2026 ~
        "Excluded from both; 2026 only",

      TRUE ~
        "Other"
    )
  ) %>%
  arrange(
    classification,
    id
  )
# representation_comparison
```

```{r}
classification_summary <- representation_comparison %>%
  count(
    classification,
    name = "number_of_anchors"
  ) %>%
  arrange(
    desc(number_of_anchors)
  )

#kable(classification_summary)
```

```{r}
final_summary <- tibble(
  category = c(
    "Retained in both 2024 and 2026",
    "Retained in 2024 only",
    "Retained in 2026 only",
    "Excluded from both; present in both raw datasets",
    "Excluded from both; present in 2024 only",
    "Excluded from both; present in 2026 only"
  ),

  number_of_anchors = c(
    nrow(retained_both),
    nrow(retained_2024_only),
    nrow(retained_2026_only),
    nrow(excluded_both_raw),
    nrow(excluded_2024_only_raw),
    nrow(excluded_2026_only_raw)
  )
)

#kable(final_summary)
```

```{r}
final_summary_percentage <- final_summary %>%
  mutate(
    "% of all observed" =
      number_of_anchors /
      length(all_raw_anchors)
  )

kable(final_summary_percentage)
```

```{r}
# redundant so exclude
# comparing 80% representation rates
representation_rates <- tibble(

  year = c(
    "2024",
    "2026"
  ),

  total_files = c(
    n_files_2024,
    n_files_2026
  ),

  eighty_percent_threshold =
    c(
      ceiling(threshold_2024),
      ceiling(threshold_2026)
    ),

  raw_anchors = c(
    nrow(all_raw_2024),
    nrow(all_raw_2026)
  ),

  retained_80pct = c(
    nrow(retained_2024),
    nrow(retained_2026)
  )
) %>%
  mutate(
    proportion_retained =
      retained_80pct / raw_anchors
  )

# representation_rates
```

```{r}
# Anchors retained in one year but not the other
year_specific_anchors <- bind_rows(

  retained_2024_only %>%
    mutate(
      classification = "Retained in 2024 only"
    ),

  retained_2026_only %>%
    mutate(
      classification = "Retained in 2026 only"
    )
) %>%
  arrange(
    classification,
    id
  )

# head(year_specific_anchors)
```

```{r}
# anchors excluded entirely but observed in either or both raw data
excluded_from_both_detailed <- representation_comparison %>%
  filter(
    classification %in% c(
      "Excluded from both; present in both raw datasets",
      "Excluded from both; 2024 only",
      "Excluded from both; 2026 only"
    )
  ) %>%
  select(
    id,
    files_2024,
    total_files_2024,
    proportion_2024,
    files_2026,
    total_files_2026,
    proportion_2026,
    classification
  ) %>%
  arrange(
    classification,
    id
  )

# head(excluded_from_both_detailed)
```

## 4.2 Most Poorly Represented Anchors

The following identifies anchors with the lowest representation in either year; only the first 6 of 65 rows are printed.

```{r}
poorly_represented <- representation_comparison %>%
  mutate(
    representation_2024 =
      replace_na(
        proportion_2024,
        0
      ),

    representation_2026 =
      replace_na(
        proportion_2026,
        0
      ),

    minimum_representation =
      pmin(
        representation_2024,
        representation_2026
      )
  ) %>%
  select(
    id,
    representation_2024,
    representation_2026,
    minimum_representation,
    classification
  ) %>%
  arrange(
    minimum_representation,
    id
  )
kable(head(poorly_represented))
```

## 4.3 Final Interpretation

1. **Stable/repeated anchors:** retained in both 2024 and 2026 analyses
2. **Year-specific anchors:** retained in one calibration period but not the other
3. **Intermittent anchors:** observed in the raw data but excluded entirely

The "Intermittent anchors" group represents anchors with consistent changes in anchor availability, connectivity, or geographic coverage. In order to reduce computational overhead and accuracy, a first step may be to exclude these anchors as input to our geolocation algorithm.

## 4.4 CSVs of Results

```{r}
write_csv(
  representation_comparison,
  "anchor_representation_comparison_all.csv"
)

write_csv(
  retained_both,
  "anchors_retained_both_2024_2026.csv"
)

write_csv(
  retained_2024_only,
  "anchors_retained_2024_only.csv"
)

write_csv(
  retained_2026_only,
  "anchors_retained_2026_only.csv"
)

write_csv(
  excluded_both,
  "anchors_excluded_both.csv"
)

write_csv(
  excluded_both_raw,
  "anchors_excluded_both_present_both_years.csv"
)

write_csv(
  excluded_2024_only_raw,
  "anchors_excluded_both_2024_only.csv"
)

write_csv(
  excluded_2026_only_raw,
  "anchors_excluded_both_2026_only.csv"
)

write_csv(
  excluded_from_both_detailed,
  "excluded_from_both_detailed.csv"
)

write_csv(
  poorly_represented,
  "poorly_represented.csv"
)
```
