---
title: "Calibration Behavior: 2024 vs. 2026"
format: pdf
editor: visual
author: "Tomoko Hida"
---
```{r, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  include = TRUE
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(cowplot)
  library(kableExtra)
})
```

```{r}
files_2024 <- list.files(
  "calibration_outputs",
  pattern = "^calibrations_probeData_2024-\\d{2}-\\d{2}\\.csv$",
  full.names = TRUE
) %>%
  sort()

len24 <- length(files_2024)
```

```{r}
raw_ids_2024 <- map_dfr(
  seq_along(files_2024),
  function(i) {

    dat <- read_csv(
      files_2024[i],
      col_names = FALSE,
      col_types = cols(.default = col_character()),
      na = c("", "NA")
    )

    tibble(
      id = as.character(dat[[1]]),
      file = basename(files_2024[i])
    )
  }
)
```

```{r}
# remove missing IDs
raw_ids_2024 <- raw_ids_2024 %>%
  filter(
    !is.na(id),
    id != ""
  ) %>%
  mutate(
    id = as.character(id)
  )
```

```{r}
n_files_2024 <- length(files_2024)
threshold_2024 <- 0.80 * n_files_2024
```

```{r}
# Count the number of files containing each anchor
anchor_counts_2024 <- raw_ids_2024 %>%
  distinct(
    id,
    file
  ) %>%
  count(
    id,
    name = "files_present"
  ) %>%
  mutate(
    total_files = n_files_2024,
    proportion_present = files_present / total_files,
    retained_80pct = files_present >= threshold_2024
  ) %>%
  arrange(
    desc(files_present),
    id
  )
```

```{r}
# anchors retained for the 2024 analysis

retained_2024 <- anchor_counts_2024 %>%
  filter(retained_80pct) %>%
  select(id) %>%
  distinct() %>%
  mutate(
    id = as.character(id)
  )

numleft24 <- nrow(retained_2024)
```
# Overview

This report compares longitudinal calibration behavior between the 2024 and 2026 calibration datasets.

The analysis reproduces the drift analysis originally applied to the 2024 calibration data. For each anchor, calibration is compared with its calibration on the first day of the observation period.

Two measures of drift are examined:

1. **Instantaneous drift:** the absolute proportional difference between an anchor's current calibration and its calibration on Day 0.
2. **Maximum drift:** the largest instantaneous drift observed up to each point in time.

The analysis also examines the number of anchors that exceed 1%, 2%, 5%, 10%, and 20% deviation thresholds.

For the 2026 data, calendar dates are preserved throughout the analysis. Thus, the `days` variable represents actual elapsed calendar days rather than the number of calibration files retained after missing days were removed.

\newpage

# 1. Load Data

The 2024 calibration data are read from individual CSV files. Each filename contains the date of retrieval, which is retained as the observation date, alongside the pre-computed calibration value.

```{r}
files_2024 <- list.files(
  "calibration_outputs",
  pattern = "^calibrations_probeData_2024-\\d{2}-\\d{2}\\.csv$",
  full.names = TRUE
) %>%
  sort()
```

```{r}
files_2024 <- list.files(
  "calibration_outputs",
  pattern = "^calibrations_probeData_2024-\\d{2}-\\d{2}\\.csv$",
  full.names = TRUE
) %>%
  sort()

# Extract dates from filenames
dates_2024 <- as.Date(
  stringr::str_extract(
    basename(files_2024),
    "\\d{4}-\\d{2}-\\d{2}"
  )
)

# Read and combine files
raw_2024 <- map_dfr(
  seq_along(files_2024),
  function(i) {

    dat <- read_csv(
      files_2024[i],
      col_names = FALSE,
      col_types = cols(
        X1 = col_character(),
        X2 = col_double()
      ),
      na = c("", "NA")
    )

    names(dat) <- c(
      "id",
      "calibration"
    )

    dat %>%
      mutate(
        date = dates_2024[i]
      ) %>%
      select(
        date,
        id,
        calibration
      )
  }
)
retained_24_calibration <- inner_join(
  retained_2024,
  raw_2024,
  by = "id"
) %>%
  arrange(date)

retained_24_calibration <- retained_24_calibration[, c(2, 1, 3:ncol(retained_24_calibration))]
```

```{r}
retained_24_calibration <- retained_24_calibration %>% 
  pivot_wider(
    names_from = "date", 
    values_from = "calibration"
  )
```

```{r}
# Keep only anchors with complete observations
complete_2024 <- retained_24_calibration[
  complete.cases(
    retained_24_calibration[, -1]
  ),
]
```

```{r}
complete_2024 <- complete_2024 %>% 
  pivot_longer(
    cols = -id,
    names_to = "date",
    values_to = "calibration"
  )

complete_2024 <- complete_2024 %>% arrange(date)
```

```{r}
# treat first calibration date as Day 0
complete_2024 <- complete_2024 %>% 
  mutate(
    date = ymd(date), 
    days = as.integer(date - min(date)) 
  )
```

```{r}
## Check date range:
# data.frame(
#   first_date = min(complete_2024$date),
#   last_date = max(complete_2024$date),
#   calendar_days = max(complete_2024$days),
#   anchors = n_distinct(complete_2024$id)
# )
```

The 2026 data requires us to calculate calibration value via given slope and intercept values.

The calibration value is calculated as:
$$
y = 100m + b
$$
where (m) is the slope and (b) is the intercept.

```{r}
files_2026 <- list.files(
  "26_Calibration",
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  sort()

len26 <- length(files_2026)
```

```{r}
df_list_2026 <- lapply(
  files_2026,
  read.csv
)

keep_2026 <- sapply(
  df_list_2026,
  function(x) {
    !is.null(x) &&
      nrow(x) >= 500
  }
)

df_list_2026 <- df_list_2026[keep_2026]

files_2026_retained <- files_2026[keep_2026]

n_files_2026 <- length(files_2026_retained)
```

```{r}
raw_ids_2026 <- map_dfr(
  seq_along(df_list_2026),
  function(i) {

    dat <- df_list_2026[[i]]

    tibble(
      id = as.character(dat[[1]]),
      file = basename(files_2026_retained[i])
    )
  }
)
```

```{r}
# Remove missing IDs:
raw_ids_2026 <- raw_ids_2026 %>%
  filter(
    !is.na(id),
    id != ""
  ) %>%
  mutate(
    id = as.character(id)
  )
```

```{r}
threshold_2026 <- 0.80 * n_files_2026
```

```{r}
# Count files containing each anchor:
anchor_counts_2026 <- raw_ids_2026 %>%
  distinct(
    id,
    file
  ) %>%
  count(
    id,
    name = "files_present"
  ) %>%
  mutate(
    total_files = n_files_2026,
    proportion_present = files_present / total_files,
    retained_80pct = files_present >= threshold_2026
  ) %>%
  arrange(
    desc(files_present),
    id
  )
```

```{r}
# The anchors retained for the 2026 analysis
retained_2026 <- anchor_counts_2026 %>%
  filter(retained_80pct) %>%
  select(id) %>%
  distinct() %>%
  mutate(
    id = as.character(id)
  )

numleft26 <- nrow(retained_2026)
```

```{r}
# locate the 2026 files
files_2026 <- list.files(
  "26_Calibration",
  pattern = "\\.csv$",
  full.names = TRUE
) %>%
  sort()

# Extract the actual calendar date from each filename
dates_2026 <- as.Date(
  stringr::str_extract(
    basename(files_2026),
    "\\d{4}-\\d{2}-\\d{2}"
  )
)

# Check files and dates
# head(data.frame(
#   file = basename(files_2026),
#   date = dates_2026
# ))
```

```{r}
# Calibration files containing fewer than 500 observations are excluded + corresponding dates removed
df_list_2026 <- lapply(
  files_2026,
  read.csv
)

keep_2026 <- sapply(
  df_list_2026,
  function(x) {
    !is.null(x) && nrow(x) >= 500
  }
)

df_list_2026 <- df_list_2026[keep_2026]
dates_2026 <- dates_2026[keep_2026]
files_2026 <- files_2026[keep_2026]

# head(data.frame(
#   file = basename(files_2026),
#   date = dates_2026
# ))
```

```{r}
# cleaning function for 2026 data

clean_df <- function(df) {

  stopifnot(ncol(df) >= 2)

  # Preserve original column names as a data row
  new_row <- as.data.frame(
    t(colnames(df)),
    stringsAsFactors = FALSE
  )

  colnames(new_row) <- colnames(df)

  df <- rbind(
    new_row,
    df
  )

  # Remove leading character from first two fields
  df[1, 1:2] <- sub(
    "^.",
    "",
    df[1, 1:2]
  )

  # Standardize column names
  df <- df %>%
    rename(
      col1 = 1,
      col2 = 2
    )

  # Split slope and intercept
  df <- separate(
    df,
    col2,
    into = c("col2", "col3"),
    sep = "b",
    fill = "right"
  )

  # keep character
  df$col1 <- suppressWarnings(
    as.character(df$col1)
  )

  df$col2 <- suppressWarnings(
    as.numeric(
      gsub(
        "[^0-9.-]",
        "",
        df$col2
      )
    )
  )

  df$col3 <- suppressWarnings(
    as.numeric(
      gsub(
        "[^0-9.-]",
        "",
        df$col3
      )
    )
  )

  # Construct calibration value
  df$col4 <- (
    df$col2 * 100 +
    df$col3
  )

  df
}
```

```{r}
# apply cleaning function
df_list_2026 <- lapply(
  df_list_2026,
  clean_df
)
```

```{r}
# keep anchors common across files, retained if it appears in at least 80% of the available calibration files
col1_list_2026 <- lapply(
  df_list_2026,
  function(df) df$col1
)

id_counts_2026 <- table(
  unlist(col1_list_2026)
)

threshold_2026 <- (
  length(df_list_2026) * 0.8
)

common_vals_2026 <- names(
  id_counts_2026[
    id_counts_2026 >= threshold_2026
  ]
)

df_list_2026 <- lapply(
  df_list_2026,
  function(df) {
    df[
      df$col1 %in% common_vals_2026,
    ]
  }
)
# 
# length(common_vals_2026)
new_df <- data.frame(
  id = unlist(common_vals_2026)
)
missing_from_retained_2026 <- new_df %>%
  anti_join(
    retained_2026,
    by = "id"
  )
```

```{r}
# build longitudinal dataset
# only anchors with complete obserrvations across retained calibration dates are used in final comparison
dfs_subset_2026 <- lapply(
  df_list_2026,
  function(d) {
    d[, c("col1", "col4")]
  }
)

dfs_subset_2026 <- Map(
  function(d, date) {

    colnames(d) <- c(
      "id",
      as.character(date)
    )

    d
  },
  dfs_subset_2026,
  dates_2026
)
```

```{r}
# combine the daily datasets
result_2026 <- Reduce(
  function(x, y) {
    merge(
      x,
      y,
      by = "id",
      all = TRUE
    )
  },
  dfs_subset_2026
)

result_2026 <- result_2026 %>%
  filter(id != "7047")
```

```{r}
# keep only anchors with complete observations
complete_2026 <- result_2026[
  complete.cases(
    result_2026[, -1]
  ),
]
```

```{r}
# convert to long format
long_2026 <- complete_2026 %>%
  pivot_longer(
    -id,
    names_to = "date",
    values_to = "calibration"
  ) %>%
  mutate(
    date = as.Date(date)
  )
```

```{r}
# calculate calendar days from the first calibration date
long_2026 <- long_2026 %>%
  group_by(id) %>%
  arrange(
    date,
    .by_group = TRUE
  ) %>%
  mutate(
    days = as.integer(
      date - min(date)
    )
  ) %>%
  ungroup()
```

```{r}
# check the resulting 2026 dataset
# data.frame(
#   first_date = min(long_2026$date),
#   last_date = max(long_2026$date),
#   calendar_days = max(long_2026$days),
#   calibration_days = n_distinct(long_2026$date),
#   anchors = n_distinct(long_2026$id)
# )
```

```{r}
# Before comparing the plots, check the size and temporal coverage of each dataset.

comparison_summary <- bind_rows(
  complete_2024 %>%
    summarise(
      year = "2024",
      anchors = n_distinct(id),
      observations = n(),
      first_date = min(date),
      last_date = max(date),
      calendar_days = max(days)
    ),

  long_2026 %>%
    summarise(
      year = "2026",
      anchors = n_distinct(id),
      observations = n(),
      first_date = min(date),
      last_date = max(date),
      calendar_days = max(days)
    )
)

kable(comparison_summary, caption = "Dataset Summaries")
```
Only anchors with complete observations across all retained calibration dates are used in the final comparison


# 2. Calculating Drift

The following calculation is equivalent to Zach's original 2024 analysis.

For each anchor:

$$
drift_t =
\frac{|calibration_t-calibration_0|}
{calibration_0}
$$
```{r}
cooked_2024 <- complete_2024 %>%
  group_by(id) %>%
  arrange(
    date,
    .by_group = TRUE
  ) %>%
  mutate(
    drift = abs(
      calibration - first(calibration)
    ) / first(calibration),

    max.drift = cummax(drift)
  ) %>%
  ungroup()

na_cooked_2024 <- cooked_2024[!complete.cases(cooked_2024), ]
cooked_2024 <- na.omit(cooked_2024)
```

```{r}
# same for 2026
cooked_2026 <- long_2026 %>%
  group_by(id) %>%
  arrange(
    date,
    .by_group = TRUE
  ) %>%
  mutate(
    drift = abs(
      calibration - first(calibration)
    ) / first(calibration),

    max.drift = cummax(drift)
  ) %>%
  ungroup()

na_cooked_2026 <- cooked_2026[!complete.cases(cooked_2026), ]
cooked_2026 <- na.omit(cooked_2026)
```

\newpage

# 3. Maximum Drift

```{r}
ggplot(
  cooked_2024,
  aes(
    x = days,
    y = max.drift,
    group = id
  )
) +
  
  # Individual anchor trajectories
  geom_step(
    alpha = 0.06,
    linewidth = 0.4,
    color = "black"
  ) +
  
  # 25th percentile
  stat_summary(
    aes(
      group = 1,
      color = "25th percentile"
    ),
    fun = \(x) quantile(x, 0.25, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # Mean
  stat_summary(
    aes(
      group = 1,
      color = "Mean"
    ),
    fun = mean,
    geom = "line",
    linewidth = 0.5
  ) +
  
  # 75th percentile
  stat_summary(
    aes(
      group = 1,
      color = "75th percentile"
    ),
    fun = \(x) quantile(x, 0.75, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 90th percentile
  stat_summary(
    aes(
      group = 1,
      color = "90th percentile"
    ),
    fun = \(x) quantile(x, 0.90, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 95th percentile
  stat_summary(
    aes(
      group = 1,
      color = "95th percentile"
    ),
    fun = \(x) quantile(x, 0.95, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 99th percentile
  stat_summary(
    aes(
      group = 1,
      color = "99th percentile"
    ),
    fun = \(x) quantile(x, 0.99, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 99.9th percentile
  stat_summary(
    aes(
      group = 1,
      color = "99.9th percentile"
    ),
    fun = \(x) quantile(x, 0.999, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # Colors and legend
  scale_color_manual(
    name = "Statistic",
    breaks = c(
      "99.9th percentile",
      "99th percentile",
      "95th percentile",
      "90th percentile",
      "75th percentile",
      "Mean",
      "25th percentile"
    ),
    values = c(
      "99.9th percentile" = "maroon",
      "99th percentile"   = "magenta",
      "95th percentile"   = "purple",
      "90th percentile"   = "blue",
      "75th percentile"   = "#20b2aa",
      "Mean"              = "red",
      "25th percentile"   = "green"
    )
  ) +
  
  labs(
    title = "2024 Maximum Calibration Drift",
    x = "Days since first calibration",
    y = "Maximum deviation from Day 0"
  ) +
  
  scale_y_continuous(
    labels = scales::percent_format(
      accuracy = 1
    )
  ) +
  
  theme_minimal() +
  
  theme(
    legend.position = "right"
  )
```
```{r}
ggplot(
  cooked_2026,
  aes(
    x = days,
    y = max.drift,
    group = id
  )
) +
  
  # Individual anchor trajectories
  geom_step(
    alpha = 0.06,
    linewidth = 0.4,
    color = "black"
  ) +
  
  # 25th percentile
  stat_summary(
    aes(
      group = 1,
      color = "25th percentile"
    ),
    fun = \(x) quantile(x, 0.25, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # Mean
  stat_summary(
    aes(
      group = 1,
      color = "Mean"
    ),
    fun = mean,
    geom = "line",
    linewidth = 0.5
  ) +
  
  # 75th percentile
  stat_summary(
    aes(
      group = 1,
      color = "75th percentile"
    ),
    fun = \(x) quantile(x, 0.75, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 90th percentile
  stat_summary(
    aes(
      group = 1,
      color = "90th percentile"
    ),
    fun = \(x) quantile(x, 0.90, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 95th percentile
  stat_summary(
    aes(
      group = 1,
      color = "95th percentile"
    ),
    fun = \(x) quantile(x, 0.95, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 99th percentile
  stat_summary(
    aes(
      group = 1,
      color = "99th percentile"
    ),
    fun = \(x) quantile(x, 0.99, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # 99.9th percentile
  stat_summary(
    aes(
      group = 1,
      color = "99.9th percentile"
    ),
    fun = \(x) quantile(x, 0.999, na.rm = TRUE),
    geom = "line",
    linewidth = 0.4
  ) +
  
  # Colors and legend
  scale_color_manual(
    name = "Statistic",
    breaks = c(
      "99.9th percentile",
      "99th percentile",
      "95th percentile",
      "90th percentile",
      "75th percentile",
      "Mean",
      "25th percentile"
    ),
    values = c(
      "99.9th percentile" = "maroon",
      "99th percentile"   = "magenta",
      "95th percentile"   = "purple",
      "90th percentile"   = "blue",
      "75th percentile"   = "#20b2aa",
      "Mean"              = "red",
      "25th percentile"   = "green"
    )
  ) +
  
  labs(
    title = "2026 Maximum Calibration Drift",
    x = "Days since first calibration",
    y = "Maximum deviation from Day 0"
  ) +
  
  scale_y_continuous(
    labels = scales::percent_format(
      accuracy = 1
    )
  ) +
  
  theme_minimal() +
  
  theme(
    legend.position = "right"
  )
```
\newpage

```{r}
# Which anchors are between 90 and 99.9 percentile for 2024?
anchors_90_999_2024 <- cooked_2024 %>%
  group_by(days) %>%
  mutate(
    q90 = quantile(max.drift, 0.90, na.rm = TRUE),
    q999 = quantile(max.drift, 0.999, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  filter(
    max.drift >= q90,
    max.drift <= q999
  ) %>%
  select(
    days,
    date,
    id,
    max.drift,
    q90,
    q999
  ) %>%
  arrange(days, desc(max.drift))
```

```{r}
# which are inbetween 90 and 99.9 percentile for 2026?
anchors_90_999_2026 <- cooked_2026 %>%
  group_by(days) %>%
  mutate(
    q90 = quantile(max.drift, 0.90, na.rm = TRUE),
    q999 = quantile(max.drift, 0.999, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  filter(
    max.drift >= q90,
    max.drift <= q999
  ) %>%
  select(
    days,
    date,
    id,
    max.drift,
    q90,
    q999
  ) %>%
  arrange(days, desc(max.drift))

# head(anchors_90_999_2024, 20)
# head(anchors_90_999_2026, 20)
```

## 3.1 90-99.9th Percentile Anchors

```{r}
ids_90_999_2024 <- anchors_90_999_2024 %>%
  distinct(id) %>%
  arrange(id)

ids_90_999_2026 <- anchors_90_999_2026 %>%
  distinct(id) %>%
  arrange(id)

ids90999num24 <- nrow(ids_90_999_2024)
ids90999num26 <- nrow(ids_90_999_2026)
```
<!-- * `r ids90999num24` of 659 anchors entered the 90-99.9th percentile range of drift in 2024 data -->
<!-- * `r ids90999num26` of 444 anchors entered the 90-99.9th percentile range of drift in 2026 data -->
Which anchors were between the 90th and 99.9th percentile on at least once?
<!-- (Code for IDs in each category is available in the qmd) -->
```{r}
persistent_90_999_2024 <- anchors_90_999_2024 %>%
  count(id, name = "days_in_range") %>%
  mutate(
    total_days = n_distinct(cooked_2024$days),
    proportion_days = days_in_range / total_days
  ) %>%
  filter(proportion_days >= 0.01) %>%
  arrange(desc(proportion_days))
```

```{r}
persistent_90_999_2026 <- anchors_90_999_2026 %>%
  count(id, name = "days_in_range") %>%
  mutate(
    total_days = n_distinct(cooked_2026$days),
    proportion_days = days_in_range / total_days
  ) %>%
  filter(proportion_days >= 0.01) %>%
  arrange(desc(proportion_days))
```
```{r}
persistent_90_999_2024 <- persistent_90_999_2024 %>%
  mutate(id = as.character(id))

persistent_90_999_2026 <- persistent_90_999_2026 %>%
  mutate(id = as.character(id))
# 
# kable(persistent_90_999_2024)
# kable(persistent_90_999_2026)
```

```{r}
persistent_90_999_both <- persistent_90_999_2024 %>%
  inner_join(
    persistent_90_999_2026,
    by = "id"
  )
# 
# kable(persistent_90_999_both)
# nrow(persistent_90_999_both)
```

```{r}
cleaned_persistent_90_999_both <- persistent_90_999_both %>%
  filter(
    !(days_in_range.x == 1 & days_in_range.y == 1)
  )
# kable(cleaned_persistent_90_999_both)
# nrow(cleaned_persistent_90_999_both)
```

* `r nrow(persistent_90_999_2024)` anchors were in the range at least once in 2024 data
* `r nrow(persistent_90_999_2026)` anchors were in the range at least once in 2026 data
* `r nrow(persistent_90_999_both)` anchors were in the range at least once in both 2024 and 2026 data
* `r nrow(persistent_90_999_both) - nrow(cleaned_persistent_90_999_both)` anchors occupied the range only one time each in both 2024 and 2026 data
* `r nrow(cleaned_persistent_90_999_both)` anchors occupied the range more than one time each between 2024 and 2026 data

```{r}
# removed_persistent_90_999_both <- persistent_90_999_both %>%
#   anti_join(
#     cleaned_persistent_90_999_both,
#     by = "id"
#   )
# 
# kable(removed_persistent_90_999_both)
```

## 3.2 Geographic Context of IDs Consistently in the 90-99.9 Range

Below, we take a look at the IDs that appeared in the 90-99.9th percentile range more than one time each between the 2024 and 2026 datasets. These anchors entered the range either once in 2024 and more than once in 2026, or more than once in 2024 and once in 2026, or entered the range more than once in both 2024 and 2026 data. We will discuss the geographic context of these `r nrow(cleaned_persistent_90_999_both)` anchors below:

```{r}
landmarks <- read_csv(
  "landmarks.csv",
  show_col_types = FALSE
) %>%
  mutate(
    pid = as.character(pid)
  ) %>%
  select(
    pid,
    asn,
    continent,
    country,
    city
  )

cleaned_persistent_90_999_both_geo <- cleaned_persistent_90_999_both %>%
  mutate(
    id = as.character(id)
  ) %>%
  left_join(
    landmarks,
    by = c("id" = "pid")
  )

# kable(cleaned_persistent_90_999_both_geo)
```
```{r}
asn_summary <- cleaned_persistent_90_999_both_geo %>%
  filter(!is.na(asn)) %>%
  count(asn, name = "anchors") %>%
# 
#   mutate(
#     percent_of_anchors = 100 * anchors / sum(anchors)
#   ) %>%
  arrange(desc(anchors), asn)

# kable(asn_summary)
```

```{r}
asn_repeated <- asn_summary %>%
  filter(anchors > 1)

kable(asn_repeated, caption = "ASN Frequency")
```
```{r}
continent_summary <- cleaned_persistent_90_999_both_geo %>%
  filter(!is.na(continent)) %>%
  count(continent, name = "anchors") %>%
  # mutate(
  #   percent_anchors = 100 * anchors / sum(anchors)
  # ) %>%
  arrange(desc(anchors), continent)

# kable(continent_summary)
continent_repeated <- continent_summary %>%
  filter(anchors > 1)

kable(continent_repeated, caption = "Continent Frequency")
```
```{r}
city_summary <- cleaned_persistent_90_999_both_geo %>%
  filter(!is.na(city)) %>%
  count(city, name = "anchors") %>%
  # mutate(
  #   percent_anchors = 100 * anchors / sum(anchors)
  # ) %>%
  arrange(desc(anchors), city)

# kable(city_summary)

repeated_cities <- city_summary %>%
  filter(anchors > 1)

kable(repeated_cities, caption = "City Frequency")
```
```{r}
country_summary <- cleaned_persistent_90_999_both_geo %>%
  filter(!is.na(country)) %>%
  count(country, name = "anchors") %>%
  # mutate(
  #   percent_anchors = 100 * anchors / sum(anchors)
  # ) %>%
  arrange(desc(anchors), country)

# kable(country_summary)

country_repeated <- country_summary %>%
  filter(anchors > 1)

kable(country_repeated, caption = "Country Frequency")
```


```{r}
# combined
# geographic_summary <- bind_rows(
#   cleaned_persistent_90_999_both_geo %>%
#     filter(!is.na(asn)) %>%
#     count(asn, name = "anchor_count") %>%
#     mutate(
#       geographic_level = "ASN",
#       location = as.character(asn)
#     ) %>%
#     select(geographic_level, location, anchor_count),
# 
#   cleaned_persistent_90_999_both_geo %>%
#     filter(!is.na(city)) %>%
#     count(city, name = "anchor_count") %>%
#     mutate(
#       geographic_level = "City",
#       location = as.character(city)
#     ) %>%
#     select(geographic_level, location, anchor_count),
# 
#   cleaned_persistent_90_999_both_geo %>%
#     filter(!is.na(country)) %>%
#     count(country, name = "anchor_count") %>%
#     mutate(
#       geographic_level = "Country",
#       location = as.character(country)
#     ) %>%
#     select(geographic_level, location, anchor_count),
# 
#   cleaned_persistent_90_999_both_geo %>%
#     filter(!is.na(continent)) %>%
#     count(continent, name = "anchor_count") %>%
#     mutate(
#       geographic_level = "Continent",
#       location = as.character(continent)
#     ) %>%
#     select(geographic_level, location, anchor_count)
# ) %>%
#   arrange(
#     geographic_level,
#     desc(anchor_count),
#     location
#   )
# 
# kable(geographic_summary)
```

```{r}
# # ASN + country +city
# 
# geo_cluster_summary <- cleaned_persistent_90_999_both_geo %>%
#   filter(
#     !is.na(asn) |
#     !is.na(country) |
#     !is.na(city)
#   ) %>%
#   count(
#     continent,
#     country,
#     city,
#     asn,
#     name = "anchor_count"
#   ) %>%
#   arrange(desc(anchor_count))
# 
# kable(geo_cluster_summary)
```


```{r}
# 
# africa_excluded <- removed_persistent_90_999_both %>%
#   mutate(
#     id = as.character(id)
#   ) %>%
#   left_join(
#     landmarks %>%
#       mutate(pid = as.character(pid)) %>%
#       select(pid, continent, country, city, asn),
#     by = c("id" = "pid")
#   ) %>%
#   filter(
#     continent == "Africa"
#   )
# 
# nrow(africa_excluded)
# kable(africa_excluded)
```

```{r}
# excluded <- removed_persistent_90_999_both %>%
#   mutate(
#     id = as.character(id)
#   ) %>%
#   left_join(
#     landmarks %>%
#       mutate(pid = as.character(pid)) %>%
#       select(pid, continent, country, city, asn),
#     by = c("id" = "pid")
#   )
# 
# nrow(excluded)
# kable(excluded)
```

\newpage

# 4. Instantaneous Drift

```{r}
ggplot(
  cooked_2024,
  aes(
    x = days,
    y = drift,
    # colour = id,
    group = id
  )
) +
  geom_step(
    alpha = 0.08,
    linewidth = 0.4
  ) +
  guides(color = "none") +
  labs(
    title = "2024 Instantaneous Calibration Drift",
    x = "Calendar days since first calibration",
    y = "Deviation from Day 0"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(
      accuracy = 1
    )
  ) +
  theme_minimal()
```

```{r}
ggplot(
  cooked_2026,
  aes(
    x = days,
    y = drift,
    # colour = id,
    group = id
  )
) +
  geom_step(
    alpha = 0.08,
    linewidth = 0.4
  ) +
  guides(color = "none") +
  labs(
    title = "2026 Instantaneous Calibration Drift",
    x = "Calendar days since first calibration",
    y = "Deviation from Day 0"
  ) +
  scale_y_continuous(
    labels = scales::percent_format(
      accuracy = 1
    )
  ) +
  theme_minimal()
```

\newpage

# 5. Maximum Drift Threshold Analysis

```{r}
thresholds <- c(
  0.01,
  0.02,
  0.05,
  0.10,
  0.20
)

thresholds.pct <- c(
  "1%",
  "2%",
  "5%",
  "10%",
  "20%"
)
```

```{r}
drift.proportion_2024 <- (
  crossing(
    cooked_2024,
    threshold = thresholds
  )
  |> group_by(
    threshold,
    days
  )
  |> summarise(
    drifted.above = sum(
      max.drift > threshold
    ),
    .groups = "drop"
  )
  |> mutate(
    threshold = ordered(
      threshold,
      levels = thresholds,
      labels = thresholds.pct
    )
  )
)
```

```{r}
ggplot(
  drift.proportion_2024,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2024$days)
    ),
    breaks = seq(
      0,
      max(cooked_2024$days),
      by = 7
    )
  ) +
  scale_y_continuous(
    "# anchors drifted past threshold",
    limits = c(
      0,
      n_distinct(cooked_2024$id)
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      "1%" = "#0868ac",
      "2%" = "#43a2ca",
      "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "2024: Anchors Ever Exceeding Drift Threshold"
  ) +
  theme_minimal()
```
```{r}
ggplot(
  drift.proportion_2024,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2024$days)
    ),
    breaks = seq(
      0,
      max(cooked_2024$days),
      by = 7
    )
  ) +
  scale_y_continuous(
    "# anchors drifted past threshold",
    limits = c(
      0,
      50
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      "1%" = "#0868ac",
      "2%" = "#43a2ca",
      "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "2024: Anchors Ever Exceeding Drift Threshold"
  ) +
  theme_minimal()
```

\newpage

```{r}
drift.proportion_2026 <- (
  crossing(
    cooked_2026,
    threshold = thresholds
  )
  |> group_by(
    threshold,
    days
  )
  |> summarise(
    drifted.above = sum(
      max.drift > threshold
    ),
    .groups = "drop"
  )
  |> mutate(
    threshold = ordered(
      threshold,
      levels = thresholds,
      labels = thresholds.pct
    )
  )
)
```

```{r}
ggplot(
  drift.proportion_2026,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2026$days)
    ),
    breaks = seq(
      0,
      max(cooked_2026$days),
      by = 7
    )
  ) +
  scale_y_continuous(
    "# anchors drifted past threshold",
    limits = c(
      0,
      n_distinct(cooked_2026$id)
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      "1%" = "#0868ac",
      "2%" = "#43a2ca",
      "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "2026: Anchors Ever Exceeding Drift Threshold"
  ) +
  theme_minimal()
```
```{r}
ggplot(
  drift.proportion_2026,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2026$days)
    ),
    breaks = seq(
      0,
      max(cooked_2026$days),
      by = 7
    )
  ) +
  scale_y_continuous(
    "# anchors drifted past threshold",
    limits = c(
      0,
      50
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      "1%" = "#0868ac",
      "2%" = "#43a2ca",
      "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "2026: Anchors Ever Exceeding Drift Threshold"
  ) +
  theme_minimal()
```


\newpage

# 6. Instantaneous Drift Threshold Analysis

This analysis counts the number of anchors that are currently beyond each threshold, rather than the number that have ever exceeded it.

```{r}
drift.proportion.2_2024 <- (
  crossing(
    cooked_2024,
    threshold = thresholds
  )
  |> group_by(
    threshold,
    days
  )
  |> summarise(
    drifted.above = sum(
      drift > threshold
    ),
    .groups = "drop"
  )
  |> mutate(
    threshold = ordered(
      threshold,
      levels = thresholds,
      labels = thresholds.pct
    )
  )
)
```

```{r}
ggplot(
  drift.proportion.2_2024,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2024$days)
    ),
    breaks = seq(
      0,
      max(cooked_2024$days),
      by = 7
    )
  ) +
  scale_y_continuous(
    "# anchors currently past threshold",
    limits = c(
      0,
      300
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      "1%" = "#0868ac",
      "2%" = "#43a2ca",
      "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "Anchors Exceeding Drift Threshold of 652"
  ) +
  theme_minimal()
```
```{r}
# closer look at 5, 10, and 20%
ggplot(
  drift.proportion.2_2024,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2024$days)
    ),
    breaks = seq(
      0,
      max(cooked_2024$days),
      by = 7
    )
  ) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), minor_breaks = seq(0, 35, by = 2.5),
    "# anchors currently past threshold",
    limits = c(
      0,
      35
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      # "1%" = "#0868ac",
      # "2%" = "#43a2ca",
       "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "Anchors Exceeding 5%, 10%, and 20% Threshold of 652"
  ) +
  theme_minimal()
```

```{r}
drift.proportion.2_2026 <- (
  crossing(
    cooked_2026,
    threshold = thresholds
  )
  |> group_by(
    threshold,
    days
  )
  |> summarise(
    drifted.above = sum(
      drift > threshold
    ),
    .groups = "drop"
  )
  |> mutate(
    threshold = ordered(
      threshold,
      levels = thresholds,
      labels = thresholds.pct
    )
  )
)
```

```{r}
ggplot(
  drift.proportion.2_2026,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2026$days)
    ),
    breaks = seq(
      0,
      max(cooked_2026$days),
      by = 7
    )
  ) +
  scale_y_continuous(
    "# anchors currently past threshold",
    limits = c(
      0,
      88
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      "1%" = "#0868ac",
      "2%" = "#43a2ca",
      "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "Anchors Exceeding Drift Threshold of 444"
  ) +
  theme_minimal()
```
```{r}
ggplot(
  drift.proportion.2_2026,
  aes(
    x = days,
    y = drifted.above,
    colour = threshold,
    group = threshold
  )
) +
  geom_step(linewidth = 1) +
  scale_x_continuous(
    "Calendar days since first calibration",
    limits = c(
      0,
      max(cooked_2026$days)
    ),
    breaks = seq(
      0,
      max(cooked_2026$days),
      by = 7
    )
  ) +
  scale_y_continuous(breaks = seq(0, 40, by = 5), minor_breaks = seq(0, 40, by = 2.5),
    "# anchors currently past threshold",
    limits = c(
      0,
      40
    )
  ) +
  scale_colour_manual(
    "Threshold",
    values = c(
      # "1%" = "#0868ac",
      # "2%" = "#43a2ca",
       "5%" = "#7bccc4",
      "10%" = "#8c96c6",
      "20%" = "#8856a7"
    )
  ) +
  labs(
    title = "Anchors 5%, 10%, and 20% Threshold of 444"
  ) +
  theme_minimal()
```
\newpage

# 7. Direct 2024 vs. 2026 Comparison

For direct comparison, it is useful to put both years onto the same plot. Because the datasets contain different numbers of anchors, percentages are more informative than raw counts.

```{r}
# maximum drift comparison
max_drift_comparison <- bind_rows(

  cooked_2024 %>%
    group_by(days) %>%
    summarise(
      anchors = n_distinct(id),
      .groups = "drop"
    ),

  cooked_2026 %>%
    group_by(days) %>%
    summarise(
      anchors = n_distinct(id),
      .groups = "drop"
    )
)
```

```{r}
# for threshold comparison

max_threshold_comparison <- bind_rows(

  crossing(
    cooked_2024,
    threshold = thresholds
  ) %>%
    group_by(
      days,
      threshold
    ) %>%
    summarise(
      drifted.above = sum(
        max.drift > threshold
      ),
      total = n_distinct(id),
      .groups = "drop"
    ) %>%
    mutate(
      year = "2024",
      proportion = drifted.above / total
    ),

  crossing(
    cooked_2026,
    threshold = thresholds
  ) %>%
    group_by(
      days,
      threshold
    ) %>%
    summarise(
      drifted.above = sum(
        max.drift > threshold
      ),
      total = n_distinct(id),
      .groups = "drop"
    ) %>%
    mutate(
      year = "2026",
      proportion = drifted.above / total
    )
) %>%
  mutate(
    threshold = factor(
      threshold,
      levels = thresholds,
      labels = thresholds.pct
    )
  )
```

```{r}
ggplot(
  max_threshold_comparison,
  aes(
    x = days,
    y = proportion,
    colour = year
  )
) +
  geom_step(linewidth = 0.5) +
  facet_wrap(
    ~ threshold,
    ncol = 5
  ) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), minor_breaks = seq(0, 1, by = 0.05),
    "Proportion of anchors",
    labels = scales::percent_format(
      accuracy = 1
    )
  ) +
  scale_colour_manual(
    values = c(
      "2024" = "#0868ac",
      "2026" = "#d95f02"
    )
  ) +
  labs(
    title = "Maximum Calibration Drift: 2024 vs. 2026",
    x = "Calendar days since first calibration",
    colour = "Year"
  ) +
  theme_minimal()
```
```{r}
# instantaneous drift comparison

instantaneous_threshold_comparison <- bind_rows(

  crossing(
    cooked_2024,
    threshold = thresholds
  ) %>%
    group_by(
      days,
      threshold
    ) %>%
    summarise(
      drifted.above = sum(
        drift > threshold
      ),
      total = n_distinct(id),
      .groups = "drop"
    ) %>%
    mutate(
      year = "2024",
      proportion = drifted.above / total
    ),

  crossing(
    cooked_2026,
    threshold = thresholds
  ) %>%
    group_by(
      days,
      threshold
    ) %>%
    summarise(
      drifted.above = sum(
        drift > threshold
      ),
      total = n_distinct(id),
      .groups = "drop"
    ) %>%
    mutate(
      year = "2026",
      proportion = drifted.above / total
    )
) %>%
  mutate(
    threshold = factor(
      threshold,
      levels = thresholds,
      labels = thresholds.pct
    )
  )
```

```{r}
ggplot(
  instantaneous_threshold_comparison,
  aes(
    x = days,
    y = proportion,
    colour = year
  )
) +
  geom_step(linewidth = 0.5) +
  facet_wrap(
    ~ threshold,
    ncol = 5
  ) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1), minor_breaks = seq(0, 1, by = 0.05),
    "Proportion of anchors",
    labels = scales::percent_format(
      accuracy = 1
    )
  ) +
  scale_colour_manual(
    values = c(
      "2024" = "#0868ac",
      "2026" = "#d95f02"
    )
  ) +
  labs(
    title = "Instantaneous Calibration Drift: 2024 vs. 2026",
    x = "Calendar days since first calibration",
    colour = "Year"
  ) +
  theme_minimal()
```

\newpage

# 8. Notes for Interpretation

The maximum-drift plots measure whether an anchor has **ever** departed from its initial calibration by a given amount. Consequently, these curves can only increase or remain constant.

The instantaneous-drift plots measure how many anchors are **currently** beyond each threshold. These curves can increase or decrease as anchors move back toward their initial calibration.

A pattern in which maximum drift increases over time while instantaneous drift remains approximately stable would suggest that individual anchors experience deviations but that the overall population does not undergo a systematic accumulation of drift.

The 2024 and 2026 comparisons should be interpreted using calendar time. Missing calibration dates in the 2026 dataset remain gaps in the time axis rather than being compressed into consecutive observation numbers.

The **last two plots** in this QMD are probably the most useful for your actual comparison. They put 2024 and 2026 on the **same percentage scale**, so the fact that you have 444 complete anchors in 2026 versus a different number in 2024 doesn't distort the comparison.


